---
title: "Comparative Study of vcd::mosaic and geom_mosaic"
output: html_document
       
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(vcd)
```



A mosaic plot is a graphical display that allows you to examine the relationship among two or more categorical variables. It is basically an area-proportional visualization of (typically, observed)
frequencies, composed of tiles (corresponding to the cells) created by recursive vertical and horizontal splits of a rectangle. Thus, the area of each tile is proportional to the corresponding cell entry given the dimensions of previous splits.
The categorical variables are first put in order. Then, each variable is assigned to an axis. In the table to the right, sequence and classification is presented for this data set. Another ordering will result in a different mosaic plot, i.e., the order of the variables is significant as for all multivariate plots.

A classic example of mosaic plots uses data from the passengers on the Titanic. The data used for this example has 2201 observations and 3 variables. The variables are:

The observations were compiled into the following table:
```{r}
str(Titanic)
```

There are several functions available to produce mosaic plot including mosaicplot
vcd::mosaic, ggmosaic

In the following sections, we are going to present a compatitive study of vcd::mosaic and ggmosaic. Out goal is descirbe the ease of the functions and some user friendly explainon of the paramters. 

We will use Titanic data to plot different mosaic plots.

vcd::mosaic:

mosaic is one of the functions of strucplot framework which we can get from importing vcd R package.The main idea of strucplot is to visualize the tables’ cells arranged in rectangular form
structable objects. The “flattened” contingency table can be obtained using the structable() function.


The basic parameters of mosaic is

mosaic(x, condvars = NULL,
  split_vertical = NULL, direction = NULL, spacing = NULL,
  spacing_args = list(), gp = NULL, expected = NULL, shade = NULL,
  highlighting = NULL, highlighting_fill = grey.colors, highlighting_direction = NULL,
  zero_size = 0.5, zero_split = FALSE, zero_shade = NULL,
  zero_gp = gpar(col = 0), panel = NULL, main = NULL, sub = NULL, ...)

OR

mosaic(formula, data, highlighting = NULL,
  ..., main = NULL, sub = NULL, subset = NULL, na.action = NULL)
 
The following is  the  description of parameters:

  
x	
a contingency table in array form, with optional category labels specified in the dimnames(x) attribute, or an object of class "structable".
x is a table or formula. structable can also be deduced from data frame using formula. For example,

```{r}
HEC <- structable(Eye ~ Sex + Hair, data = HairEyeColor)
HEC
```



formula	
a formula specifying the variables used to create a contingency table from data. For convenience, conditioning formulas can be specified; the conditioning variables will then be used first for splitting. If any, a specified response variable will be highlighted in the cells.

For example, If we want to split the plot by 1 varialbe, we can specify as
~V1
If V2 is dependent on V1 and needs to be highlighted, we can do
V2 ~ V1

In the following example, a mosaic plot is crated from Titanic data:

```{r}
mosaic(Survived ~ Sex + Age, data = Titanic,
main = "Survival on the Titanic")
```

In the above plot, the data is partinioned by sex first and then age; after that survival is used to show the dependency
```{r}
mosaic(Survived ~ ., data = Titanic)
```



This type of formuall is used when one variable is needed to be highlighted based on rest of the varialbe. Survived is highlihgted based on all otther dependend variables.

data	
either a data frame, or an object of class "table" or "ftable".

If we provide, structable contingency table, we do not need to provide data (df or table)


direction	
character vector of length k, where k is the number of margins of x (values are recycled as needed). For each component, a value of "h" indicates that the tile(s) of the corresponding dimension should be split horizontally, whereas "v" indicates vertical split(s).

By default, vcd::mosaic() splits horizontally from left for the first variable (e.g. ~ V1, ~ V1 + V2, ~ V2 | V1 or V2 ~ V1 -- here V1 is considered as first variable) and vertically from top for the second variable and so on. If we want to alter this direction, we can provide this "direction" parameter. For example,

```{r}
 mosaic(Survived ~ Sex + Age, data = Titanic, main = "Survival on the Titanic",direction=c("v","v","h"))
```
 
 # Sex = Vertical, Age = Vertical, Survived = Horizonal
 In the above plot, the sex variable is used to split the plot vertically, then age variable is used to sub-devide verticaly and then survived sub-division is highlighted horizontally from other side.
 
gp	
object of class "gpar", shading function or a corresponding generating function (see details and shadings). Components of "gpar" objects are recycled as needed along the last splitting dimension. Ignored if shade = FALSE.

if gp is not provided, the default gray color is applied with shade = TRUE

```{r}
mosaic(Survived~ Sex + Age, data = Titanic,
       main = "Survival on the Titanic", shade = TRUE,
       direction=c("v","v","h"))
```

if gp is provided with shade = TRUE (or no shade parameter but not shade = FALSE), the provided color along with other graphical paramters (if provided any) will apply.


```{r}
mosaic(Survived~ Sex + Age, data = Titanic,
       main = "Survival on the Titanic", shade = TRUE,
       direction=c("v","v","h"), gp = gpar(fill=c('lightblue', 'pink'))
                 )
```
                 
shade	
logical specifying whether gp should be used or not (see gp). If TRUE and expected is unspecified, a default model is fitted: if condvars (see strucplot) is specified, a corresponding conditional independence model, and else the total independence model.

If we provide shade = FALSe, the gp will not apply at all.


```{r}
mosaic(Survived~ Sex + Age, data = Titanic,
       main = "Survival on the Titanic", shade = FALSE,
       direction=c("v","v","h"), gp = gpar(fill=c('lightblue', 'pink'))
                 )
```
                 




Reference:
1) https://pdfs.semanticscholar.org/2416/2c1a46669f94356854176ece8548bf7fb989.pdf
2) https://ncss-wpengine.netdna-ssl.com/wp-content/themes/ncss/pdf/Procedures/NCSS/Mosaic_Plots.pdf
3) https://en.wikipedia.org/wiki/Mosaic_plot
4) http://www.pmean.com/definitions/mosaic.htm


